---
title: "Homework 2"
author: Mufeng Xu
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx" ,
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017) 

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The 
dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.


## Problem 2

Import and Clean NYC Transit Dataset

```{r}
transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% #select the variables of interest
  mutate(
    entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE),
    ) #change entry class from character to logical#clean names for column titles
```


The dataset contains variables for line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. After importing the dataset by `read_csv`, its variable names were cleaned by `janitor::clean_names`. Then, I selected the variables of interest by using `select`. Also, `recode` were used to switch the entry variable from character to logical. In the resulting dataset, there are `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. So far, the dataset is not tidy because the route variables are spreading across 11 columns. 

`distinct` function was used to calculate the number of distinct stations. It was found that there are `r transit_df %>% distinct(line, station_name) %>% nrow` distinct stations. Out of these station, there are `r transit_df %>% filter(ada == TRUE) %>% distinct(line, station_name) %>% nrow()` stations that are ADA compliant. The proportion of station entrances/exits without vending allow entrance is `r (transit_df %>% filter(entry == "TRUE", vending =="NO") %>% nrow / transit_df %>% filter(vending == "NO") %>% nrow())*100`%.


Reformat the dataset further more
```{r}
transit_tidy = transit_df %>% 
  mutate_at(
    vars(route8:route11),
    as.character
    ) %>% #change class from double to character
  
  pivot_longer(
    route1:route11,
    names_prefix = "route",
    names_to = "route_num",
    values_to = "route_name"
    ) %>% #reformat route number and route name so that they are distinct variables
  relocate(route_num, route_name) %>% # place column route_num, and route_name to the front
  arrange(route_num)
```


There are `r transit_tidy %>% filter(route_name == "A") %>% distinct(station_name, line) %>% nrow()` distinct stations that serve the A train. Of the stations that serve the A train, `r transit_tidy %>% filter(route_name == "A") %>% distinct(station_name, line, ada) %>% filter(ada == "TRUE") %>% nrow()` are ADA compliant.

## Problem 3



